\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-03-24 ase20/intro.tex
[ASE] revised intro

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/existing.png}
\vspace*{-2em}
  \caption{Existing approaches: Manual extraction of IR-based semantics for ECMAScript}
  \label{fig:existing}
\vspace*{-1em}
\end{figure}

Researchers have defined various JavaScript formal
semantics~\cite{aplas08,lambdajs,kjs,javert} and developed several static
analyzers on top of them~\cite{jsai,tajs,wala,safe}.  As illustrated
in Figure~\ref{fig:existing}, they defined their own Intermediate
Representations (IRs) to represent JavaScript program semantics.  They
built parsers to construct Abstract Syntax Trees (ASTs) from
programs and AST-IR translators to convert ASTs
to their own IRs.  This approach helps researchers focus on IRs
without worrying about diverse and enormous features of JavaScript in
developing new techniques for program analysis and formal verification.
We dub such a methodology that manipulates language semantics by
translating given programs to IRs that are suitable for analysis and
verification \textit{IR-based semantics extraction}.

However, existing approaches to JavaScript IR-based semantics
extraction have difficulties in manipulating recent JavaScript
programs, because they all extract the semantics \textit{manually}.
The JavaScript syntax and semantics are described in the ECMAScript
specification, written in English in a clearly structured and
organized manner.  Unfortunately, manually implementing parsers and
translators for JavaScript from the specification is tedious,
labor-intensive, and error-prone especially because the language is
huge.  The most recent specification, ECMAScript 2019, has 798 pages.

Moreover, manual extraction does not scale for frequent updates of the
specification.  Most existing formal semantics and analyzers of
JavaScript support only ECMAScript 5.1~\cite{ecma5} released in June 2011.
Focusing on ECMAScript 5.1 was reasonable until 2014, because the
specification was rarely updated.  However, in late 2014, The Ecma
Technical Committee 39 (TC39)~\cite{tc39} decided to release the
specification annually.  Currently, no formal semantics
nor analyzers for JavaScript correctly deal with recent ECMAScript
features such as lexical binding via \( \code{let} \), the spread
\( \code{...} \) operator, classes, the \( \code{for-of} \) operator,
the \( \code{async} \) functions, and generators.

In this paper, we propose JavaScript IR-based Semantics Extraction
Toolchain (\( \tool \)), which \textit{automatically} extracts JavaScript
syntax and semantics from the ECMAScript specification.  It
automatically generates a JavaScript parser from the syntax defined in
the ECMAScript specification.  It also automatically extracts the
formal semantics of JavaScript with given \textit{compile rules}.
Each compile rule describes how abstract algorithms in the
specification written in a natural language are converted into
functions of our intermediate representation \( \ires \).  In
addition, to save the labor of defining such compile rules, we also
present a \textsf{Fault Localizer}, which suggests new
compile rules based on fault localization of not yet convertible
algorithms.  It helps us deal with changed or newly
introduced algorithms defined in new versions of the ECMAScript
specification as well.

The main contribution of this paper is \( \tool \):
\begin{itemize}[leftmargin=0.5cm]
\item We successfully applied \( \tool \) to ECMAScript 2020, the next
ECMAScript version scheduled to be released in June 2020.  The
automatically extracted semantics passed the official conformance test
suite, Test262~\cite{test262}; out of 19,282 tests for
ECMAScript 2020 it passed 19,220 tests.
\item \( \tool \) is also applicable to new language features proposed
for a future release to aid language development in a modular way.
Because ECMAScript is an open standard language, various proposals are
available with their own specification changes and tests.  As a case
study, we applied \( \tool \) to such a proposal, during which the tool
found some errors in the proposal.  The reported errors were all confirmed
and the tool passed all the tests.
\item Moreover, since the extraction mechanism of \( \tool \) is fully
automatic, we can use it to find possible specification errors.  We
indeed found 3 specification errors in ECMAScript 2020.  They were all
confirmed by TC39 and will be fixed in the next release.
\end{itemize}

>>>>

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/existing.pdf}
\vspace*{-2em}
  \caption{Existing approaches: Manually implemented parsers and AST-IR
  translators for IR-based semantics of JavaScript programs}
  \label{fig:existing}
\vspace*{-1em}
\end{figure}

Researchers have defined various JavaScript formal
semantics~\cite{aplas08,lambdajs,kjs,javert} suitable for static
analysis~\cite{jsai,tajs,wala,safe} and formal verification~\cite{javert} by
referring to ECMAScript specification.  ECMAScript specification officially
describes JavaScript syntax using a variant of extended BNF notation, and its
semantics using abstract algorithms written in English in a clearly and
structured manner.  As illustrated in Figure~\ref{fig:existing}, they built
parsers that constructs Abstract Syntax Trees (ASTs) of given JavaScript
programs, and AST-IR translators to convert ASTs to their own Intermediate
Representations (IRs). This approach helps researchers focus on IRs without
worrying about diverse and enormous features of JavaScript in developing new
techniques for program analysis and formal verification.  We dub such
methodology that manipulates language semantics by translating given programs to
IRs \textit{IR-based semantics extraction}.

However, to the best of our knowledge, all existing approaches to JavaScript
IR-based semantics extraction \textit{manually} build parsers and translators.
The manual approach is tedious, labor-intensive, and error-prone.  For example,
KJS~\cite{kjs} is one of formal semantics for traditional JavaScript described
in ECMAScript 5.1 specification~\cite{ecma5} (ES5.1, 2011) and it is defined on top of \(
\kframework \), which is a framework for defining language semantics.  According
to an author of KJS, it took \textit{four months} to implement AST-IR translator
for 1,370 lines out of 2,932 lines in 368 abstract algorithms. However, modern
JavaScript becomes much more massive than traditional one and the most recent
version of ECMAScript specification (ES10, 2019) has 1,731 abstract algorithms
consisting of 10,068 lines.  It means that it is not scalable to build AST-IR
translator for modern JavaScript.  Thus, most of JavaScript formal semantics
targets for ES5.1 and no formal semantics deals with modern JavaScript features
such as lexical binding via \( \code{let} \), the spread \( \code{...} \)
operator, classes, the \( \code{for-of} \) operator, the \( \code{async} \)
functions, and generators.

Moreover, JavaScript syntax and semantics is annually updated.  Until ECMAScript
5.1 specification, JavaScript is quite stable language because the specification
has been rarely updated.  However, The Ecma Technical Committee 39
(TC39)~\cite{tc39} decided to release the specification annually in late 2014.
The first specification (ES6, 2015) after this official announcement includes
various modern JavaScript languages features. In every year, several syntax and
roughly 1,000 to 3,000 lines of abstract algorithms are modified or newly added
in the specification. Thus, we need to update JavaScript parser every year while
existing work does not care about parser itself because the stable parser for
traditional JavaScript (ES5.1) already existed. Besides, the manual approach
does not scale for AST-IR translators to deal with this frequent and massive
updates of ECMAScript specification.

To alleviate this problem, we proposed a technique to \textit{automatically
synthesize} parsers and AST-IR translators directly from ECMAScript
specifications. There are several technical challenges in synthesizing parsers
and translators. For syntax, ECMAScript specifications utilize their own special
variant of extended BNF with parametric non-terminals, conditional
alternatives, and various special terminal symbols. Thus, no existing parser
generation technique is directly applicable for this variant. Moreover,
JavaScript provides automatic semicolon insertion in parsing algorithm not in
lexer with several complex rules. For semantics, abstract algorithms in
ECMAScript specifications are written in English and no reference interpreter to
check the synthesized AST-IR translators for future specifications.  Moreover,
the general representation of abstract algorithms is necessary for the forward
compatibility of our technique.

Our contribution is \( \tool \), a JavaScript IR-based Semantics Extraction
Toolchain:
\begin{itemize}[leftmargin=0.5cm]
  \item \( \tool \) \textit{first} automatically synthesizes parsers and AST-IR
    translators for modern JavaScript directly from ECMAScript specifications.
    For syntax, we first formalize \( \bnfes \), the variant of extended BNF
    used in the specification, and propose the parser generation technique for
    \( \bnfes \) with automatic semicolon insertion.  While parsers are
    generated in a fully automatic way, synthesis of AST-IR translators are
    semi-automatic assisted by \textit{compile rules} because of the difficulty
    to deal with semantics written in natural language. Each compile rule
    describes how each line of abstract algorithms is converted into our
    intermediate representation \( \ires \) designed for ECMAScript
    specifications.
  \item We first define formal semantics of modern JavaScript via \( \tool \).
    We applied our tool to the next version of specification (ES11) scheduled to
    be released in June 2020.  The synthesized parser and translator passed the
    official conformance test suite, Test262~\cite{test262}; out of
    \inred{XX,XXX} applicable tests for ES11 it passed \inred{XX,XXX}
    tests.  Moreover, we found a 3 specification errors in ES11 by
    using the synthesized parsers and translators.  They were all confirmed by
    TC39 and will be fixed in the next release.
  \item \( \tool \) is also adaptable for new language features will be proposed
    in future ECMAScript specifications.  TC39 manages proposals for new
    language features in an open source
    project\footnote{https://github.com/tc39/proposals}. As a case study, we
    applied \( \tool \) to feature proposals in Stage 3 and we found some errors
    in the language feature \inred{XXX} and \inred{XXX}. The reported errors
    were all confirmed as real bugs.
\end{itemize}
